Enrich the datasets with new columns.

### 2020-08-25 inst/analysis/analysis_stable.Rmd
With tags

# This makes it possible to generate the html file in the html subfolder
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "html") })
---

@@

library(tidyverse)
library(stringr)
library(viridis)
library(vctrs)
library(DT)
library(fst)
library(fs)

@@

source("inc/paths.R")
source("inc/latextags.R")
create_tags(path(TAGS_DIR, "analysisStable.tex"), prefix="", default=TRUE)

data_filepath <- EVALS_SUM_PKGS_FILE
#data_filepath <- "/var/lib/R/project-evalR/run/package-evals-traced-all.1/calls.fst"
# The trace data on all CRAN does not have all the columns of the smaller corpus. The analysis will fail for the big corpus.
```

@@

The analysis is being performed on dataset read from file `r data_filepath`.

```{r file-info, echo=FALSE}
file_info(data_filepath)
```

```{r load-data, echo=FALSE}
eval_calls <-
    read_fst(data_filepath) %>%
    as_tibble() %>%
    mutate(eval_call_package = extract_package_name(eval_call_srcref, file))

@@

```{r parse_in_base}
parsed_base <- eval_calls %>% filter(!is.na(expr_parsed_expression), eval_call_package == "base")

@@

There are `r count(parsed_base, wt = nb_ev_calls)` `eval` calls with `parse` in *base*, which is `r count(parsed_base, wt = nb_ev_calls) / count(filter(eval_calls, eval_call_package == "base"), wt=nb_ev_calls)`%.

@@

@@

@@

@@

expr_resolved_monomorphic <- expr_resolved_polymorphism %>% filter(polym_degree == 1)
nb_monomorphic <- expr_resolved_monomorphic %>% nrow()
```

```{r polymorphism}
expr_resolved_polymorphism %>% ggplot() +
  geom_bar(aes(x = polym_degree)) +
  scale_x_continuous(breaks=c(0, 1, 5, 10)) +
  labs(title = "Degree of polymorphism", x = "degree", y = "n") 
```

@@

```{r eval_from_parse}
parsed_evals <- eval_calls %>% filter(!is.na(expr_parsed_expression))
nb_parsed_evals <- parsed_evals %>% count(wt = nb_ev_calls)
```

@@

```{r, echo = FALSE}
nb_parsed_call_sites <- nb_eval_call_sites(parsed_evals)
```

@@

  We can differentiate further the provenance, i.e. whether the evaluated string comes from `file` or from `text`. However, we can only do it for `parse`, `str2lang` and `str2expression` at the top level of `expr_expression`, not buried in a function call.

@@

>>>>

# This makes it possible to generate the html file in the html subfolder
# knit: (function(inputFile, encoding) {
#   rmarkdown::render(inputFile, encoding = encoding, output_dir = "html") })
---

@@

library(tidyverse)
library(stringr)
library(viridis)
library(vctrs)
library(DT)
library(fst)
library(fs)
library(ggthemes)

@@

source("inc/paths.R")
source("inc/latextags.R")
source("inc/setup.R")
theme_set(theme_minimal())
create_tags(path(TAGS_DIR, "analysisStable.tex"), prefix="", default=TRUE)

packages_filepath <- EVALS_SUM_PKGS_FILE
core_filepath <- EVALS_SUM_CORE_FILE
```

@@

The analysis is being performed on dataset read from files `r packages_filepath` and `core_filepath`.

```{r file-info-packages, echo=FALSE}
file_info(packages_filepath)
```

```{r file-info-core, echo=FALSE}
file_info(core_filepath)
```

```{r load-data, echo=FALSE}
eval_calls_packages <-
    read_fst(packages_filepath) %>%
    as_tibble()  %>% 
    mutate(dataset = "packages")

eval_call_core <- 
  read_fst(packages_filepath) %>%
  as_tibble() %>%
  mutate(dataset = "core")

eval_calls <- bind_rows(eval_calls_packages, eval_call_core)

@@

```{r parse_in_base}
parsed_base <- eval_calls %>% filter(!is.na(expr_parsed_expression), eval_source == "base")

@@

There are `r count(parsed_base, wt = nb_ev_calls)` `eval` calls with `parse` in *base*, which is `r count(parsed_base, wt = nb_ev_calls) / count(filter(eval_calls, eval_source == "base"), wt=nb_ev_calls)`%.

@@

### String size

@@

### AST size

```{r}
ast_sizes <- eval_calls %>% slice_sample(n=10000) %>% mutate(expr_resolved_ast_size = map_int(expr_resolved, expr_size_str))
```

```{r}
# ast_sizes %>% ggplot(aes(x = expr_resolved_ast_size, weight = nb_ev_calls, group = dataset)) +
#   geom_histogram(aes(y = stat(width*density)), na.rm = TRUE, binwidth = 0.1) +
#   scale_x_log10() +  scale_y_sqrt(labels=scales::percent) +
#   labs(x = "AST size", y = "relative frequencies")

# ast_sizes %>% 
#   ggplot(aes(x = dataset, y = expr_resolved_ast_size, fill = dataset, weight = nb_ev_calls)) +
#   geom_flat_violin(trim=FALSE, width=.5) +
#   geom_dotplot(
#     binaxis="y", 
#     dotsize=0.01, 
#     stackdir="down", 
#     binwidth=0.01, 
#     position=position_nudge(-0.025)) +
#   scale_y_log10(labels = scales::comma) +
#   labs(y="AST size (log)") +
#   theme(
#     legend.box="horizontal",
#     axis.title.x=element_blank()
#   )

ast_sizes %>%
  ggplot(aes(x = dataset, y = expr_resolved_ast_size, fill = dataset, weight = nb_ev_calls)) +
  geom_violin(trim = FALSE, width=.5)+
  geom_boxplot(width=.05) +
  scale_y_log10(labels = scales::comma) +
  labs(y="AST size (log)") +
  theme(
    legend.box="horizontal",
    axis.title.x=element_blank()
  )

ggsave(path(PLOT_DIR, "ast_sizes.pdf"))
```

@@

One of the main differences of the R \eval with the javascript one is how it interacts with \emph{scope}, or environments in R. In Javascript, \c{eval} can access the local scope and the global scope. In R, \c{eval} has an argument \c{envir} to indicate in which environment its first argument, after being resolved, must be evaluated.

By default, it is equal to \c{parent.frame()}, which is the parent environment of the \c{eval} call.

```{r default_environment}
default_envir <- eval_calls %>% filter(is.na(envir_expression)) 
nb_default_envir <- default_envir %>% count(wt = nb_ev_calls)
nb_call_sites_default_envir <- nb_eval_call_sites(default_envir)
```

`r ratio(nb_default_envir, nb_eval_calls)`% of the calls use the default environment, which corresponds to `r ratio(nb_call_sites_default_envir, nb_call_sites)`% of the call sites.

We also classify environments in the following way:

- `base`: environment of a primitive function of *base* package
- `empty`: empty environment
- `callee`: same environment (it seems it only happens with `local` )
- `global`: ` .GlobalEnv` 
- `package:package_name`: package environment
- `loop`: there was a loop in the frame

```{r environment_hierarchy}
# classified_envir <- eval_calls %>% filter(!is.na(envir_expression)) %>%
#       mutate(env_class = pmap(list(environment_class, envir_type, envir_expression), extract_envir)) %>% 
#       unnest_wider(env_class)

#TODO
```

@@

expr_resolved_monomorphic <- expr_resolved_polymorphism %>% filter(polym_degree == 1)
nb_monomorphic <- expr_resolved_monomorphic %>% nrow()

r("nb monomorphic", nb_monomorphic)
r("percent monomorphic", ratio(nb_monomorphic, nb_call_sites))
```

```{r polymorphism}
expr_resolved_polymorphism %>% ggplot() +
  geom_bar(aes(x = polym_degree)) +
  scale_x_continuous(breaks=c(0, 1, 5, 10)) +
  labs(title = "Degree of polymorphism", x = "degree", y = "n")

ggsave(path(PLOT_DIR, "polymorphism.pdf"))
```

@@

```{r eval_from_parse}
parsed_evals <- eval_calls %>% filter(!is.na(expr_parsed_expression))
nb_parsed_evals <- parsed_evals %>% count(wt = nb_ev_calls)

r("percent parsed evals",  ratio(nb_parsed_evals, nb_eval_calls))
```

@@

```{r, echo = FALSE}
nb_parsed_call_sites <- nb_eval_call_sites(parsed_evals)

r("percent parsed call sites", ratio(nb_parsed_call_sites, nb_call_sites))
```

@@

  We can differentiate further the provenance, i.e. whether the evaluated string comes from `file` or from `text`.

@@

We also look to the occurences of `paste` and `str_c` in the `parse` argument. It is an indication that meta-pogramming happens.

```{r parse_meta_prog}
meta_parsed <- parsed_texts %>% filter(str_detect(expr_parsed_expression, "paste|str_c"))

nb_meta_parsed <- meta_parsed %>% count(wt = nb_ev_calls)
nb_meta_parsed_call_sites <- nb_eval_call_sites(meta_parsed)

r("paste parsed evals", ratio(nb_meta_parsed, nb_parsed_evals))
r("paste parsed call site", ratio(nb_meta_parsed_call_sites, nb_parsed_call_sites))
```

### 2020-08-25 inst/analysis/enrich_dataset.Rmd
With tags

---
title: "Enrich the datasets with new columns."
output:
  rmdformats::readthedown:
        lightbox: true
        gallery: false
        df_print: paged
        toc_depth: 3
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
params:
  base_dir: /var/lib/R/project-evalR
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "html") })
---

@@

library(tidyverse)
library(stringr)
library(viridis)
library(vctrs)
library(DT)
library(fst)
library(fs)

@@

>>>>

---
title: "Enrich the datasets with new columns."
output: html_document
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
params:
  base_dir: /var/lib/R/project-evalR
---

@@

library(tidyverse)
library(stringr)
library(viridis)
library(vctrs)
library(DT)
library(fst)
library(fs)
library(parallel)

@@

```{r ast_size}
eval_calls <- eval_calls %>% add_ast_size()
```

```{r add_package}
eval_calls <- eval_calls %>% add_package()
```
